Baseline centering problem

### 2018-12-21 paper/tangotd.tex

Namely, set
\begin{equation}
Z=\transp{\green}F, \qquad \green=\sum_s \green(s)\transp{\1_s},\qquad F=\E_{s\sim
\rho}
f(s)\transp{\1_s}
\end{equation}
which is the same representation as above, namely $Z=\Delta^{-1}$ is
equivalent to $\green(s)\cdot \E_{s'\sim
\rho} \psi(s')f(s')=\transp{\1_s}\Delta^{-1}\psi$ for all $s$ and test
functions $\psi$. Then the Newton equation for $Z$ reads
\begin{equation}
\transp{\green}F\gets (1+\eta)\transp{\green}F-\eta \transp{\green}F \E_{ss'} \frac{1}{\rho(s)}\1_s\transp{(\1_s-\gamma
\1_{s'})} \transp{\green}F
\end{equation}
and since $F\1_s=\rho(s)f(s)$ and
$\transp{\1_s}\transp{\green}=\transp{\green(s)}$, this simplifies to
\begin{equation}
\transp{\green}F\gets (1+\eta)\transp{\green}F-\eta
\transp{\green} \E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}F
\end{equation}
and we propose to update separately for $\green$ given $f$ and $f$ given
$\green$: the advantage being that for fixed $f$, we can obtain the
update above by updating $\green$ via
\begin{equation}
\transp{\green}\gets (1+\eta)\transp{\green}-\eta
\transp{\green} \E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}
\end{equation}
and likewise for $f$ given $\green$,
\begin{equation}
F\gets (1+\eta)F-\eta
\E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}F
\end{equation}
both of which can be implemented via gradient descent on the parameters
of $\green$ and $f$. (This double the learning rate $\eta$.)
There is no need to sample on test functions.

@@

\paragraph{Better variant?}
Everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem; still, it looks suspiciously like $TD(1)$ except maybe for the
$O(1/(1-\gamma))$ variance). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.

>>>>

Namely, set
\begin{equation}
Z=\transp{\green}F, \qquad \green=\sum_s \green(s)\transp{\1_s},\qquad F=\E_{s\sim
\rho}
f(s)\transp{\1_s}
\end{equation}
which is the same representation as above, namely $Z=\Delta^{-1}$ is
equivalent to $\green(s)\cdot \E_{s'\sim
\rho} \psi(s')f(s')=\transp{\1_s}\Delta^{-1}\psi$ for all $s$ and test
functions $\psi$. Then the Newton equation for $Z$ reads
\begin{equation}
\transp{\green}F\gets (1+\eta)\transp{\green}F-\eta \transp{\green}F \E_{ss'} \frac{1}{\rho(s)}\1_s\transp{(\1_s-\gamma
\1_{s'})} \transp{\green}F
\end{equation}
and since $F\1_s=\rho(s)f(s)$ and
$\transp{\1_s}\transp{\green}=\transp{\green(s)}$, this simplifies to
\begin{equation}
\transp{\green}F\gets (1+\eta)\transp{\green}F-\eta
\transp{\green} \E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}F
\end{equation}
and we propose to update separately for $\green$ given $f$ and $f$ given
$\green$: the advantage being that for fixed $f$, we can obtain the
update above by updating $\green$ via
\begin{equation}
\transp{\green}\gets (1+\eta)\transp{\green}-\eta
\transp{\green} \E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}
\end{equation}
and likewise for $f$ given $\green$,
\begin{equation}
F\gets (1+\eta)F-\eta
\E_{ss'} f(s)\transp{(\green(s)-\gamma
\green(s'))}F
\end{equation}
both of which can be implemented via gradient descent on the parameters
of $\green$ and $f$. (This doubles the learning rate $\eta$.)
There is no need to sample on test functions.

@@

\paragraph{Better variant?}
Everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
[TODO CHANGE NOTATION because currently the entries of $Z$ are not
$Z(s_1,s_2)$ but $Z_{s_1s_2}=\rho(s_2)Z(s_1,s_2)$ which is confusing...]
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem; still, it looks suspiciously like $TD(1)$ except maybe for the
$O(1/(1-\gamma))$ variance). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.
